---
title: 'Brain Eno'
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    storyboard: true
---

```{r, cache = FALSE}
library(tidyverse)
library(plotly)
library(spotifyr)
library(dplyr)
library(ggplot2)
source(file='spotify.R')
```

### Introduction
The man you probably heard of, either as synth man in Roxy music or as U2 producer. Brain Eno has certainly left his hark in the music industry. Not only has he produced very many albums, he even created a new music genre; ambient.

Since he has his own ways of producing (giving cards with random words on them to the U2 band members, to tell them how to sing and play). I decided to compare his solo work with his work as a producer, to see what his influences are in the work of albums he produced for certain bands. He produced music for several artists, but I am going to focus on Coldplay, U2, Devo, David Bowie, Genesis and the Talking Head. Because I think these bands had the most commercial success with their music produced under Eno.

I created two playlists. One of his solowork, consisting of only solo albums, ranging from art-/experimental rock-ish music, to ambient (to which he now still sticks). And one of alle the albums he produced with the aformentioned artists/bands.

### Characteristics

```{r, echo=FALSE}
enoproducer <- get_playlist_audio_features('21sm2glgseqq6otym3iwfbwsa', '0TEl4LaLPMjr7qPta5Z9pW')
enosolo <- get_playlist_audio_features('21sm2glgseqq6otym3iwfbwsa', '57ZXLIpoMu7mY8jWlWnCQl')

enoproducer %>% summarise(M = mean(valence), SD = sd(valence))
enosolo %>% summarise(M = mean(valence), SD = sd(valence))

enoproducer %>% summarise(mean = mean(energy), SD = sd(energy))
enosolo %>% summarise(M = mean(energy), SD = sd(energy))
```
When first analysing the music, I decided to take a look at valence and energy.
It is very striking that the valence and energy of Eno's solo work (0.23 and 0.24) are a lot lower than that of his work as a producer (0.47 and 0.66).

### Findings
```{r}
enoproducer %>% ggplot(aes(x = valence, y = energy)) + geom_point() + geom_smooth() + ggtitle('Eno producer')
enosolo %>% ggplot(aes(x = valence, y = energy)) + geom_point() + geom_smooth() + ggtitle('Eno solo')
```
***
You can see these details visualised here. Every point represents a song. You can see that a lot of songs from his solo work are concentrated in the bottom left corner of the graph. Meaning low valence and energy (also known as 'sad' songs). Most songs he produces are found in the upper half of the graph, meaning they are eiter 'angry' (top left corner) or 'happy' (top right corner).

